---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r Loading Data}
activities <- read.table(unz("activity.zip","activity.csv"),head=TRUE,sep=",",na.strings ="NA")
head(activities)
```

## What is mean total number of steps taken per day?
```{r Summarize Data}
require(dplyr)
totalStepsPerDay <- activities %>% group_by(date) %>% summarize(steps=sum(steps,na.rm=TRUE))
hist(totalStepsPerDay$steps,xlab="Total Steps Per Day",main="Histogram for Total Steps Per Day")
mean(totalStepsPerDay$steps,na.rm=TRUE)
median(totalStepsPerDay$steps,na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r Daily Average}
require(dplyr)
require(ggplot2)
averageStepsPerInterval <- activities %>% group_by(interval) %>% summarize(mean=mean(steps,na.rm=TRUE))
averageStepsPerInterval$interval <- as.factor( sprintf("%02d:%02d",averageStepsPerInterval$interval%/%100,averageStepsPerInterval$interval%%100))
ggplot(averageStepsPerInterval,aes(interval,mean)) + 
          geom_line(group=1) + 
          scale_x_discrete(breaks=c("00:00","06:00","12:00","18:00","23:55")) +
          labs(x="Time of Day",y="Daily Steps Average") +
          geom_vline(xintercept = which.max(averageStepsPerInterval$mean), colour = "red", linetype="dotted" )
```

#### Interval with the Maximum Mean
```{r Max Mean Interval}
averageStepsPerInterval[which.max(averageStepsPerInterval$mean),]
```

## Imputing missing values
#### Total Number of Rows with NA values
```{r}
nrow(activities[is.na(activities$steps),])
```
### Create new data set and replace NAs with interval median
#### Compute for the median per interval.  COMPUTING PER DAY WILL NOT YIELD ANY DIFFERENCE..
```{r}
activities$steps <- ifelse(is.na(activities$steps),
                           ave(activities$steps,activities$interval,FUN=function(x) median(x,na.rm=TRUE)),
                           activities$steps)
```

#### Create new report after replacing NAs with computed values
```{r Summarize New Data}
newTotalStepsPerDay <- activities %>% group_by(date) %>% summarize(steps=sum(steps))
hist(newTotalStepsPerDay$steps,xlab="New Total Steps Per Day",main="New Histogram for Total Steps Per Day")
mean(newTotalStepsPerDay$steps,na.rm=TRUE)
median(newTotalStepsPerDay$steps,na.rm=TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?
### Add a column indicating the day of the week for the dataset
```{r}
activities$wday <- as.factor(  ifelse( as.POSIXlt(activities$date)$wday %in% c(0,6),
                                       "Weekend",
                                       "Weekday"
                                     )
                            )
head(activities)

averageStepsPerInterval <- activities %>% group_by(interval,wday) %>% summarize(mean=mean(steps,na.rm=TRUE))
averageStepsPerInterval$interval <- as.factor( sprintf("%02d:%02d",averageStepsPerInterval$interval%/%100,averageStepsPerInterval$interval%%100))
ggplot(averageStepsPerInterval,aes(interval,mean)) + 
          geom_line(group=1) + 
          scale_x_discrete(breaks=c("00:00","06:00","12:00","18:00","23:55")) +
          labs(x="Time of Day",y="Daily Steps Average") +
          facet_grid( wday ~ .)

```
